Summary Statistics

### 2025-01-07 extensive_tests/draft-sumstat-vignette.Rmd

We calculate the frequency of sites that changes for each structure.

@@

NOTE: Summary statistics related with tree cherries will require a

>>>>

We calculate the frequency of sites that changes for each structure (island and non-island structures).

Let $S_k$ be the set of sites in the cherry structure of interest, where $k$ represents the index of the cherry structure. For each site in $S_k$, let the methylation state at a given tip $t$ be denoted as $m_{k,s,t}$, where $s \in S_k$ is the index of the site, and $m_{k,s,t} \in {0, 0.5, 1}$ represents the methylation state at tip $t$ (unmethylated, partially-methylated, or methylated, respectively).

For a given cherry structure, we compute in the proportion of sites that have a different methylation state between two tips, say $t_1$ and $t_2$.

Define the indicator variable:
\[
\delta_{k,s,t_1,t_2} =
\begin{cases}
1, & \text{if } m_{k,s,t_1} \neq m_{k,s,t_2}, \\
0, & \text{if } m_{k,s,t_1} = m_{k,s,t_2}.
\end{cases}
\]

Then, the summary statistic for the proportion of sites with different methylation states between tips $t_1$ and $t_2$ in the cherry structure $k$ is:
\[
\text{Proportion}_k(t_1, t_2) = \frac{1}{|S_k|} \sum_{s \in S_k} \delta_{k,s,t_1,t_2}
\]

where $|S_k|$ is the total number of sites in the cherry structure $k$, and the sum counts the number of sites with different methylation states between tips $t_1$ and $t_2$.

This statistic gives the proportion of sites in the cherry structure $k$ where the methylation state differs between the two tips of interest.

@@

NOTE: Summary statistics related with tree cherries require a tree with tip labels as:
